---
title: "SMA-RILES"
author: "Pablo Busch"
date: "3/22/2022"
output:
  html_document:
    df_print: paged
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.pos = 'H', 
                      fig.width=12, fig.height=8,
                      message=F, echo = F, warning=F)
```

# Introduction

This document provides an analysis on the data for RILES (water emissions from Industrial Sources or "Residuos liquidos industriales", available to download here [https://snifa.sma.gob.cl/DatosAbiertos]). \newline

The data is reported at UF (Unidad Fiscalizable) level, which consist on the basic unit of the SMA (Superintendencia del Medio Ambiente) to identify common approved projects associated with the same company.


```{r}
# Load Riles Data and preparation
source("Scripts/00-Common.R", encoding = "UTF-8")
source("Scripts/00-Functions_Spatial.R",encoding = "UTF-8")
emisiones <- read_rds("Data/SMA/emisiones_RILES.rds") 
```

# Data Exploration

The data contains the following columns of interest:

- Razon Social: "Nombre del titular de la fuente de emisión que reporta". We have more than 40 companies in the zone of interest.
- Planta: "Nombre de la fuente de emisión que reporta". A few sources have more than two plants, with different locations (lat/long).
- Punto de Descarga: Same as planta. "Nombre del punto de descarga de la fuente emisora"
- Norma: Always **D.S. 90.**
- Tipo de Control: Always **Autocontrol**
- Laboratorio: "Nombre del laboratorio que toma las muestras". Around 8 different laboratories.
- UnidadFiscalizable: Cross to project data contained in other dataset: UF
- Parametro: "Parámetro contaminante reportado"
- Unidad: "Unidad del parámetro reportado"
- Valor reportado: "Emisión reportada"
- Caudal Muestra (m3/dia): Flow of the reported value of "muestra".

Based on this information, it seems that we can compile monthly emissions at the level of **Planta**, which represent a point source where emissions occur to a water body. \newline

## Data conversion

As most parameters are measured as concentration in the sample (e.g. mg/L), we use the daily reported water flow measured in $m^3$/day to estimation monthly emissions as:

$$
Montly \: Emissions = Daily \: Flow [\frac{m^3}{day}] * Concentration \: normalized [\frac{kg}{m^3}]*Number \:of \: days 
$$

Where concentration normalized is simply the concentration expressed in terms of kg per $m^3$. Note, for parameters dimensionless, like T° or pH, we can only aim to report a weighted (by water flow) average per month.

# Analysis

## Number of projects with Water Emissions 

```{r}

# number of projects per commune
n_fuentes <- emisiones %>% 
  group_by(RegionNombre,ComunaNombre,NombreCategoria,RazonSocial,Planta) %>% 
  tally()

n_fuentes$geometry <- NULL


# of projects per commune
n_projects <- n_fuentes %>% ungroup() %>% 
  group_by(RegionNombre,ComunaNombre,NombreCategoria) %>% 
  tally() %>% 
  pivot_wider(names_from = NombreCategoria, values_from = n) %>% 
  # mutate(sum = rowSums(across(where(is.numeric)))) %>% # sum of columns: Total
  janitor::adorn_totals("row") %>% 
  rename(Region=RegionNombre,Comuna=ComunaNombre)

n_projects %>% 
  flextable() %>% 
  autofit() %>% 
  merge_v(j=1) %>% 
  hline(i=c(2,7,11), part="body") %>% 
  bold(i=12) %>% bold(part="header")

rm(n_fuentes,n_projects)
```

We observe that most of the projects with declared water emissions correspond to fishing and aquaculture. Lather in this document, a map is provided to see the location of each project.


## Example: Nitrógeno Total  Kjeldahl

Let's estimate the water emissions for one parameter: Total Kjeldahl Nitrogen. As the same plant could have multiple samples we first take the average concentration and average water flow.

```{r}
# Emisions work ------

emisiones_n <- emisiones %>% 
  filter(Parametro=="Nitrógeno Total  Kjeldahl") %>% 
  group_by(PeriodoInforme, RazonSocial, Parametro,
           NombreCategoria,Planta,PuntoDeDescarga,
           RegionNombre,ComunaNombre) %>% 
  summarise(Valor_reportado=mean(Valor_reportado),
            `Caudal_Muestra_(m3/dia)`=mean(`Caudal_Muestra_(m3/dia)`)) %>% ungroup()


# data conversion mg/L to kg/m3
emisiones_n <- emisiones_n %>% 
  mutate(conc_norm=Valor_reportado/1e6*1e3,
         emision=`Caudal_Muestra_(m3/dia)`*conc_norm*31) # in kg


emisiones_n$NombreCategoria <- factor(emisiones_n$NombreCategoria)
```

Let's check at the total emissions for 2020:

```{r}
emisiones_n %>%
  mutate(Mes=month(PeriodoInforme) %>% as.factor) %>% 
  mutate(emision=emision/1e3) %>% 
 ggplot(aes(reorder(Planta, emision),emision, fill=Mes))+
  geom_col()+
  theme(legend.position = "")+
  coord_flip(expand = F)+
  labs(x=" Planta",y="Total Kjeldahl Nitrogen [ton]",
       caption = "Colors represented monthly emissions")
```


## Map

First a map of all sources that have water emissions in the zone.

```{r, fig.width=9, fig.height=6}
# MAP ----

# Summarize info of each emission source
emisiones_fuentes <- emisiones %>% 
  group_by(NombreCategoria,RazonSocial,Planta,PuntoDeDescarga,
           RegionNombre,ComunaNombre,CodigoRETC) %>% 
  summarise(Parametro=paste0(unique(Parametro),collapse="; "))

# add line very 40 characters
emisiones_fuentes <- emisiones_fuentes %>% 
  mutate(Parametro=paste(strwrap(Parametro,width=100),
                         collapse ="<br/>"))
# interactive map
m_fuentes <- mapview(emisiones_fuentes, zcol=c("NombreCategoria"))
m_fuentes

# mapshot(m_fuentes, "Figures/Maps/UF_Riles.html", selfcontained=F)

```

Let's look at a map with the emissions sources for Total Kjeldahl Nitrogen:

```{r, fig.width=9, fig.height=6}
Emisiones_Nitrogeno_2020 <- emisiones_n %>% 
    group_by(NombreCategoria,RazonSocial,Planta,
             PuntoDeDescarga,RegionNombre,ComunaNombre) %>% 
  summarise(emision=sum(emision)) %>% ungroup() %>% 
  mutate(emision=round(emision/1e3,2)) %>% 
  rename(`Emisiones Nitrogeno Kjeldhal [ton]`=emision)

m_emisiones <- mapview(Emisiones_Nitrogeno_2020, 
        zcol=c("NombreCategoria"),
        cex="Emisiones Nitrogeno Kjeldhal [ton]")
m_emisiones

# mapshot(m_emisiones, "Figures/Maps/Emisiones_Riles.html", selfcontained=F)

```




